---
title: "Expanded Research Design"
author: "Luke Sheppard"
institute: "University of Pennsylvania"

date: last-modified
toc: true

bibliography: references.bib


format: 
  html: 
    self-contained: true

editor: source
project:
  output-dir: "C:/Users/lukes/OneDrive/Documents/Github/psci3200_lukesheppard"
---
1.  Introduction
    
    Budget support vs. Project Aid: Government have information advantages but subject to massive leakage (Rinikka & Svensson, 2004). Debate on accoutnability and risk of fungibility and donor vs. recipient control. 
    Aid Distribution on implementation through NGOs and Government affecting instiutional capacity based on percieved corruption by recipient country and geopolitical importance (donor foreign policy objectives and ignorance) 
    Multilateral vs. vilateral donors
    Commitments ratehr than disbursements
    Project Aid must be proximate, local providing of goods/services
    Country Fixed Effects
    Lagged IV to DV
    Control for obvious pre-treament confounders
    Think about "most likely" design? Such that in worse case scenarios effect not observed or more likely in best case scenarios effect observed. 
    
    Effect of bypass aid: aid circumventing state institutions by relying on NGOs, effect on state capacity? History of regressive aid projects after structural readjustment from IMF loan forgiveness plans. Citizen attitudes on NGOs contribute to government distrut. NGO service delivery weaken "Fiscal Link" (reliance on taxation, PSD, weaken link between performance and citizen welfare) and undermines accountability (politicians get credit for NGO PSD (Phillipines article) and create apolitical attitudes about government). NGOs contribute primarily to health, education, agriculture, and social infrastructure costs moreso than government institutions, donors prefer NGOs over local capabilities. NGOs are unableb to contribute to industries with hardcore infrastructure such as transportation, energy, water/sanitation, business/trade development. How does this impact citizen perception of government service delivery, governments primarily provide high cost infrastructure rather than direct delivery service. 
    
    Scale (conceptual complications in increasing project to larger population) and implementation (shifting from NGO to government provision) difficulties of projects.
    
    US foreign aid obitruary: Health, democracy and governance, and soft power
    NGO restriction laws targeting NGOs for democracy/governance promotion rather than health or education aid. Coutries wary of foreign interference nuance from Western imperialism. 
    Soft power: Does aid win hearts and minds. Aid improves health/education outcomes, recipients seem to know who gives aid, aid often comes with conditionality and corruption, does aid improve trust in US and European institutions. 

    I plan to investigate whether US foreign aid projects fuel local-level (petty) or elite corruption. I am interested in this question in light of current events and the resulting restructuring of US foreign aid efforts. Additionally, this question addresses the continuing debate on the effectiveness of foreign aid, especially on recipient institutional capacity. Although estimates of aid lost to corruption are widely overstated, according to Charles Kenny from the Center for Global Development, there remains the threat that aid funding is mismanaged and misappropriated by bad state actors creating inefficient or ineffective programs and further regressive distribution. However, by nature aid flows to countries with lower institutional capacities and high levels of corruption, thus some mismanagement of foreign aid is inevitable. In this project, I plan to estimate if foreign aid exacerbates existing corruption i.e. there exists a causal relationship between foreign aid deliverance and institutional corruption. Existing literature highlights the possibility of regressive effects due to foreign aid. In a 2020 study, the Development Research Group of the World Bank Group found that aid disbursements correspond with investments in tax safe havens, elite corruption, suggesting an aid leakage rate around 7.5%. Additionally, in an experiment on Uganda education transfers, significant leakage was observed by local officials varying strongly by community bargaining power. Thus a progressive program exacerbated regressive outcomes as corruption affected aid disbursements along existing income lines. Similarly, qualitative and quantitative geocoded analysis from Ryan Briggs, "Why Does Aid Not Target the Poorest?" displays how World Bank aid traditionally targets richer parts of communities despite higher returns in poorer, more remote areas due to implementation difficulties. Foreign aid leakage from local and elite corruption has distributional consequences and may guide resources and state capacity towards wealthier communities. Finally, governance aid projects, designed to build state capacity and public perception has mixed effects as seen in a analysis by Development Impact Evaluation Group within the World Bank Group preferring instead to strengthen political incentives, communication networks and other community mechanisms to support local trust and legitimacy. In summary, foreign aid projects are rife to corruption and thus incentive projects in wealthier, more stable regions thereby decreasing efficiency and distributing limited state capacity in a regressive manner. I plan to analyze US aid projects by category: humanitarian, economic, military, and development, and development sectors: education, governance, healthcare, infrastructure, etc. with perceptions of corruption in a cross country analysis. Most research on foreign aid and corruption focuses on multilateral aid, specifically the World Bank or UN, yet bilateral aid encompasses a significant portion of global aid often with less oversight and coordination. Although humanitarian and military aid would be expected to target areas specifically in crisis, economic and development aid would be expected to target countries where poverty prevention and long-term political, social, and economic development seem feasible i.e. sufficient state capacity to implement projects partnering with recipients. I am ultimately interested in whether these bilateral development aid (projects not initiated as part of multilateral development or for strategic reasons) are effective or fuel local-level (petty) or elite corruption.

2.  Hypotheses

    "Bilateral development utilizes less partners" as multiple stakeholders mean aid must flow to places where preferences overlap and bilateral lacks this and has 
    Although foreign aid is increasingly effective and resistance to leakage, bilateral development utilizes less partners, less multilateral oversight, and is dependent on domestic factors. As a result, I predict bilateral development aid has variable results which vary sharply by sector. Specifically, I predict that aid which delivers services such as healthcare, infrastructure, and education which are reliably susceptible to leakage and corruption will improve perceptions of corruption while governance aid, aid specifically designed to promote state capacity, will have limited effects. Thus increasing US aid funding for sectors delivering services will cause decreased perceptions of corruption. In comparison, increasing or decreasing US aid funding for governance projects will cause negligible changes in perceptions of corruption. These predictions reflect earlier cited research and my own theoretical reasoning that citizens will respond to high implementation projects due to service delivery and the bias towards implementation of these projects in wealthier areas. Inversely, governance projects targeting unstable areas will suffer from existing corruption, target areas with traditional implementation problems, and have minimal impact on citizenry. Finally, military and humanitarian aid are highly realizable services in emergency areas. Thus, increasing US aid funding for military and humanitarian aid will cause decreased perceptions of corruption, despite intrinsically replacing state development and capacity.

3.  Data

    Ultimately, local or elite corruption is incredibly difficult to measure accurately. Inherently obscured and protected, any measure of corruption relies on significant assumptions. Effective measures to observe aid leakage and state corruption have been designed but for this project I am more interested in perceptions of local and state corruption as effective benchmarks and as governance tools in and of themselves. World Bank Group research has shown that citizen perceptions of state trust and legitimacy directly effect state capacity and developmental outcomes. However, citizen perceptions are volatile and likely to under or overestimate real levels of corruption. To better define causality, I will utilize geocoded data and a difference-in-difference approach which should protect against exogenous factors that affect citizen perceptions irrespective of foreign aid. For the independent variable, I will use AidData's Geocoded Research Releases on Development Projects covering many countries including Malawi, Somalia, Bangladesh, DRC, Nigeria, Iraq, Colombia, Nepal, Uganda, Senegal, Honduras, Timor-Leste, CAR, Burundi, and Sierra Leone. This data captures development projects by donor, recipient, finances, sector, and geocoded location over similar years (until 2014). The unit of analysis will be project and I will filter by donor for US projects and sector for different regressions. For the dependent variable I will use different barometers including Geocoded Afrobarometer and Latinobarometro providing local interviews across time on state legitimacy and corruption as measured by a variety of variables including administrative capacity ("the perceived performance of local and national institutions in solving specific problems, such as the provision of basic education and health services"), political involvement ("levels of civic engagement and political participation"), state perception ("engagement with and confidence in local and central government institutions), administrative capacity (the physical presence or absence of state institutions (e.g. police stations, health clinics, schools, water and sanitation systems)"), local capacity ("the reported quality of local public services"), and corruption ("local perceptions of and experiences with various forms of corruption"). While these variables are similar, they highlight important differences such as local v central state capacity and faith in institutions, all over time. The unit of analysis is individuals.
    
```{r}
#| echo: false
#| message: false
#| warning: false
#| include: true
#| eval: true
#| setup
```

```{r}
    #| echo: true
    #latino <- Latinobarometro_2023_Eng_v1_0
    library(haven)
    latino <- read_dta("C:/Users/lukes/OneDrive/Documents/GitHub/psci3200_lukesheppard/Latinobarometro_2023_Eng_Stata_v1_0.dta")
    table(latino$idenpa)
    colombia <- subset(latino, idenpa == 152)
    #Measure of elite corruption: P12ST. Generally speaking, would you say that (country) is governed by a few powerful groups in their own interest? Or is it governed for the good of all?
    table(colombia$P12ST)
    mean(colombia$P12ST)
    #Measure of national government trust
    table(colombia$P13ST_E)
    mean(colombia$P13ST_E)
    #Measure of perception of corruption
    table(colombia$P60ST)
    mean(colombia$P60ST)

    projects <- read.csv("C:/Users/lukes/OneDrive/Documents/GitHub/psci3200_lukesheppard/ColombiaAIMS_GeocodedResearchRelease_Level1_v1.1.1/data/projects.csv")
    location <- read.csv("C:/Users/lukes/OneDrive/Documents/GitHub/psci3200_lukesheppard/ColombiaAIMS_GeocodedResearchRelease_Level1_v1.1.1/data/locations.csv")
    projects <- subset(projects, donors_iso3 == "USA")
    colombia.aid <- merge(projects, location, by = "project_id")

    #ggplot() + 
      #geom_line(data=colombia, aes(x=year, y=P60ST), color='blue') + 
      #geom_line(data=colombia.aid, aes(x=transactions_start_year, y=total_commitments), color='red')"

    #ggplot(colombia, aes(x=colombia.aid, y=P60ST)) + geom_point() +
      #labs(y="Perception of Corruption", x = "Aid Projects")

    #Using different corruption measures, I would plot their value every year with aid projects in a line graph. Further analysis would be done by mapping location ID from aid projects and region/city variable in latinobarometro. I would also include total aid and then seperated by sector, dependent on having sufficient observations of a sector (Colombia as an example has fewer aid projects). Significant data cleaning still required.

    #I would repeat the process for each country listed above with geocoded aid project data provided by AidData
    ```

4.  Regression Design
  a. Regression Model

```{r}
#| echo: false
#lm(P60ST ~ colombia.aid, colombia)
#Renamed corruption variable by number or funds of project
#Final regression will be 
#lm(Corruption ~ Aid + country.fixed.effects or region.fixed.effects + year.fixed.effects + individual.controls)

#https://doi.org/10.1016/j.jpubeco.2018.01.002
#Emulating a research design used on Chinese foreign aid in Africa but at the region/city level (minmizes causality but far better than country data and managemable)
#lm(Corruption ~ Active.aid.dummy.variable + Inactive.aid.dummy.variable + country.fixed.effects or region.fixed.effects + year.fixed.effects + individual.controls)
#No aid project being omitted variable
```

$$
Y_i = \beta_1 \times active_i + \beta_1 \times inactive_i + \alpha_s + \theta_t + \lambda \times X_i + \epsilon
$$

  b. Empirical Extension